A weighted image reconstruction based on PCA for pedestrian detection

### 2010-05-07 paper/paper.tex
Adição do abstract e correções gerais.

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{Investigating PCA image reconstruction for pedestrian detection}

@@

\iffinal
  \author{
     \IEEEauthorblockN{Guilherme V. Carvalho, Lailson B. Moraes, George D. C. Cavalcanti, Tsang I. R.}
     \IEEEauthorblockA{Centro de Informática (CIn)\\
      Universidade Federal de Pernambuco (UFPE)\\
      Recife, Pernambuco, Brasil\\
      \texttt{\url{http://cin.ufpe.br/~viisar}}\\
      \texttt{\url{{gvc,lbm4,gdcc,tir}@cin.ufpe.br}}}

@@

\begin{abstract}
The abstract goes here. DO NOT USE SPECIAL CHARACTERS, SYMBOLS, OR MATH IN YOUR TITLE OR ABSTRACT, AND I'M NOT SCREAMING!. Lorem ipsum dolor sit amet, consectetur adipisicing elit, sed do eiusmod tempor incididunt ut labore et dolore magna aliqua. Ut enim ad minim veniam, quis nostrud exercitation ullamco laboris nisi ut aliquip ex ea commodo consequat. Duis aute irure dolor in reprehenderit in voluptate velit esse cillum dolore eu fugiat nulla pariatur. Excepteur sint occaecat cupidatat non proident, sunt in culpa qui officia deserunt mollit anim id est laborum.

\end{abstract}

\begin{IEEEkeywords}
TODO; component; formatting; style; styling;
\end{IEEEkeywords}

@@

Pedestrian detection systems have been widely used and developed throughout the computer vision history. Ranging from still image detection to automated car breaking systems, it is becoming a usual task in our lives.

In this paper we describe a pedestrian detection proposed by Malangon-Borja \cite{borja09}. It classifies an image based on how much its reconstruction is different from the original one, that is, it classifies based on the image reconstruction error. We analyze how it works and how its original performance can be improved.

The proposed improvements enhances the system's accuracy using weights on the detection errors and significantly reduces execution times by decreasing the number of calculations made by the algorithm.

% and, by using weights on the reconstruction errors, we improve the system's accuracy.

This paper is divided as follows. Section 2 describes an image reconstruction technique using PCA. Section 3 describes how to use image reconstruction to classify images. Section 4 presents our experiments and discussions. In section 5 conclusions are presented and we indicate areas of further research.

@@

The idea of principal component analysis (PCA) seems to have been first proposed by Pearson in 1901 \cite{sirovich87} and was later developed in many works during the last century. It was popularized in computer vision by Turk and Pentland, that in 1991 used PCA to propose the eigenfaces \cite{turk91}, a face recognition method still relevant and used in current days.

Given a set of samples, PCA yields a set of orthonormal vectors that can be used to linearly project the samples into a new space. This space maximizes the projected samples variance and minimizes their least mean square error, that is, it minimizes the difference between the projected sample and its reconstruction back to the original space \cite{shlens09}. This characteristic makes PCA suitable for data compression and dimensionality reduction. The data can be anything, but in this work we are concerned only with images.

The formulation of PCA is as follows \cite{borja09}. Consider a set of $m$ images, each with $r$ pixels of width and $c$ pixels of height (i.e. of size $r \times c$ pixels). Each image $I_i$ is represented by a column vector $v_i$ of length $rc$. The set mean $\mu$ and the covariance matrix $C$ are given by

@@

PCA applies eigen-decomposition on $C$ and keeps the $k$ eigenvectors ($1 <= k <= rc$) corresponding to the $k$ largest eigenvalues. These eigenvectors are called the \emph{principal components} (PCs) of $C$. It is proven that a projection onto the space defined by these eigenvectors provides the optimal reduced representation of the data, minimizing information loss \cite{shlens09}.

@@

\noindent As the projection is reversible, we can try to recover the original version of a previously projected image. The \emph{reconstructed image} $u'$ of a projection $p$ is

@@

\noindent This process can be interpreted as form of lossy compression and decompression: the reconstructed image is generally not equal to the original one, being just an approximation of it. For this reason, each reconstruction has an associated \emph{reconstruction error}, which can be measured by

@@

In general, the more principal components we use to obtain a projection, the less information loss we have, what allows a more accurate reconstruction. Also, the more similar $u$ is to the images that were used to produce $P$, the better the reconstruction will be for a fixed number of eigenvectors.

@@

Based on this fact, Malagón-Borja proposed a classifier for pedestrian detection that uses reconstruction errors as the classification criteria \cite{borja09}. In addition to grayscale images, the classifier also use the corresponding computed edge images, since they characterize better the pedestrian silhouette and eliminate background variations.

@@

\begin{itemize}
	\item The principal components $P_{gp}$ and the mean $\mu_{gp}$ from the set of pedestrian grayscale images.
	\item The principal components $P_{ep}$ and the mean $\mu_{ep}$ from the set of pedestrian edge images.
	\item The principal components $P_{gn}$ and the mean $\mu_{gn}$ from the set of non-pedestrian grayscale images.
	\item The principal components $P_{en}$ and the mean $\mu_{en}$ from the set of non-pedestrian edge images.
\end{itemize}

In the classification phase, when a new grayscale image has to be classified, the first step is obtain its edge image and perform four reconstructions: one for each set of PCs. Then, the reconstruction errors must be calculated and combined to produce the total reconstruction error, that is the final classification score. If the total error is greater than or equal to zero, the image is classified as a pedestrian; otherwise, it is classifier as a non-pedestrian. The detailed procedure is described above.

\begin{enumerate}
  \item Obtain the edge image $e$ from grayscale image $g$ using the Sobel operator.
  \item Perform four image reconstructions:
    \begin{enumerate}
      \item $r_{gp} = P_{gp} P_{gp}^T(g-u) + \mu_{gp}$
      \item $r_{ep} = P_{ep} P_{ep}^T(e-u) + \mu_{ep}$
      \item $r_{gn} = P_{gn} P_{gn}^T(g-u) + \mu_{gn}$
      \item $r_{en} = P_{en} P_{en}^T(e-u) + \mu_{en}$
    \end{enumerate}
  \item Calculate the corresponding reconstruction errors:
    \begin{enumerate}
      \item $d_{gp} = |r_{gp} - g|$
      \item $d_{ep} = |r_{ep} - e|$
      \item $d_{gn} = |r_{gp} - g|$
      \item $d_{en} = |r_{en} - e|$
    \end{enumerate}
  \item Compute the \emph{total reconstruction error}, given by
    \begin{equation}
      d_t = d_{gn} + d_{en} - d_{gp} - d_{ep}
      \label{total_error}
    \end{equation}
  \item Classify the image according to:
    \begin{equation}
      \textrm{class}(g) = \left\{
        \begin{array}{l l}
         \textrm{Pedestrian}, & d_t >= 0\\
         \textrm{Non-pedestrian}, & d_t < 0
        \end{array}
      \right.
    \end{equation}

@@

When we analyze how this classifier works, it is easy to understand the role of the PCs generated by the positive samples. As the samples contains the recurrent pattern of the target object (pedestrians in this case), PCA is able to capture it and better reconstruct similar images, generating smaller reconstruction errors. The role of negative PCs was obscure, though. Since the negative images contain just random objects and does not represent a specific concept, what pattern is there to capture? We initially thought that the errors of reconstructions obtained from the negative PCs had a negligible importance and therefore it would be possible to eliminate them without impacting the classification accuracy. However, our experiments show that is not the case: it turns out that these errors have a major importance on classification.

We found that the value of the negative PCs is not on the pattern PCA captures, but instead it lies on the pattern PCA does \emph{not} capture. In this way, the negative reconstruction errors tend to be high for images that contains the target object while for the same images the positive reconstruction errors tend to be small, contributing to a high total error, what classifies the image correctly. For images that does not contains the target object, it is exactly the opposite. Thus, both components of total error (from positive PCs and from negative PCs) work together for classification and that is why it is not feasible to construct a classifier that uses only positive reconstruction errors.

However, this does not mean that all the four components have the same relevance to classification. In the reconstruction error classifier, each component contributes equally to the total error, but our experiments show that this relation is not true. Some components have more importance than others and therefore have to be adjusted accordingly. So, we propose the \emph{weighted total reconstruction error}, given by

@@

It is also worth noting that the classifier can be made more flexible if we introduce a threshold parameter and compare the classification score (i.e. the total error) agains it, instead of keep it always fixed to zero. In this way, the user can adjust the specificity, depending on his particular application.

@@

The classifier was tested in the context of the pedestrian detection task, that is, given an unlabeled image, the classifier must determine if it depicts or not a pedestrian. To train the classifier, we need a set of pedestrians images and a set of non-pedestrian images. This training phase consists of computing image edges and applying PCA to the four image sets separately, what produces four sets of PCs. They will be used to perform the reconstructions on classification time.

The pedestrian images were obtained from the MIT CBCL pedestrian database \cite{cbcl}. It contains 924 color images of pedestrians in frontal or rear views, under different scene conditions. These images were converted to grayscale and had part of the background cropped on the border, remaining with $45 \times 105$ pixels. Doing this, we reduce the non-relevant background variation among the images, what allows a better classification.

@@

To train the classifier, we selected 75\% of the pedestrian images (693 samples) and the remaining images (231 samples) were used  for testing; from the non-pedestrian images, we selected 80\% of the images (4.000 samples) for training the classifier and the remaining images (1.000 samples) were used for testing. This particular scheme was chosen in a empirical way; it yielded the best results among the distributions we tested. The edge images are computed with $x$ and $y$ Sobel filters \cite{gonzalez01}, which we combine to generate a single edge image. To implement the classifier and the auxiliary procedures, we used the MATLAB 2009b environment. The resulting code is available for download at \url{http://github.com/lailsonbm/pca_reconstruction/}.

We start our analysis by investigating the impact of reconstruction errors from the positive PCs and from the negative PCs to the classification. As discussed before, we initially questioned the role of negative PCs, since they are computed from completely uncorrelated images and do not represent a particular object. However, when we classify test samples using reconstruction errors only from the positive PCs or only from the negative PCs (but still using grayscale and edge images), it becomes clear that neither one can perform a satisfactory classification; both classifiers yield a poor result. But when the errors are combined accordingly to the total reconstruction error, we have a good accuracy. The receiver operating characteristic (ROC) curves of the three classifiers are shown in Figure~\ref{roc_pos_vs_neg} for $k = 100$. We get a similar pattern for all other values of $k$ that were tested.

\begin{figure}[b!]
\centering
\includegraphics[width=3in]{roc_pos_vs_neg}
\caption{Comparison of positive and negative ROC curves. From left to right: the first curve (green) refers to a classifier that uses the total reconstruction error (\ref{total_error}) and has an AUC of 0.995; the second curve (red) refers to a classifier that uses only the reconstruction errors from negative PCs ($d_{gn}$ and $d_{en}$) and has an AUC of 0.690; and the third curve (blue) refers to a classifier that uses only the errors from positive PCs ($d_{gp}$ and $d_{ep}$) and has an AUC of 0.444. On the three classifiers, only the first 100 principal components are being used, that is, $k=100$. Both positive and negative classifiers yields a poor result, far from the full classifier performance. It is surprising to see that the negative classifier even got a better result than the positive one.}
\label{roc_pos_vs_neg}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=3in]{roc_gray_vs_edge}
\caption{Comparison of grayscale and edge ROC curves. The green curve refers to a classifier that uses the total reconstruction error (\ref{total_error}) and has an AUC of 0.995; the red curve refers to a classifier that uses only the reconstruction errors from edge PCs ($d_{ep}$ and $d_{en}$) and has an AUC of 0.994; and the blue curve uses only the errors from grayscale PCs ($d_{gp}$ and $d_{gn}$) and has an AUC of 0.985. On the three classifiers, only the first 100 principal components are being used, that is, $k=100$. Note that the scale is different of that from Figure~\ref{roc_pos_vs_neg}; we approximated the upper left corner to better show the small difference between the curves.}
\label{roc_gray_vs_edge}
\end{figure}

Again, we plot the reconstruction error distributions, but now segregating errors from grayscale PCs and from edges PCs. These plots are shown in Figure~\ref{dist_gray_vs_edge}, that also displays a plot of the total reconstruction error. Now, we get a different scenario. All the three plots exhibit a clear boundary that separate samples from different classes. This indicate that there is a lot of redundancy between errors from grayscale PCs and from edge PCs and one of them can be safely removed from the classifier with just a minor impact on classification accuracy. Our results indicate that the errors from grayscale PCs generally yields a slightly worse result, as shown in the Table~\ref{table_gray_vs_edge}. Therefore, it is possible to compose a classifier only with the edge errors and speed up the classification significantly (since two reconstructions will not have to be computed anymore). In general, such classifier drops the classification accuracy in less than $1\%$ when compared with a classifier that uses the total error. Actually, the classifier based only on the edge errors sometimes performs even better than the full classifier in our experiments (when $k >= 200$), as also shown in Table~\ref{table_gray_vs_edge}.

@@

% What reference use to GA?
These findings point out that, among the four reconstruction errors, some are more relevant to the classification than others. However, on the total reconstruction error (\ref{total_error}) each error contributes equally. So, we use the weighted total reconstruction error (\ref{weighted_total_error}) to improve the classification accuracy. The problem with using weights is how to choose them. We can try some values and make adjusts manually, test after test. But we can also employ an optimization algorithm that looks for the values in an automated way. In this work, we use a genetic algorithm \ref{ga}. It initially chooses a set of random values and uses operations as crossover and mutation to improve them during many iterations. These operations are inspired by evolutionary biology; hence the name genetic algorithms. The values are adjusted in relation to a fitness function. In our case, this functions aims to correctly classify the largest number of training samples using a given set of weights. The function can be fine-tuned in order to give more importance in finding more pedestrians despite the false detections or to minimize the number of non-pedestrians classified as pedestrians.

As this algorithm performs a local optimization and therefore is not optimum, we runned it five times for each value of $k$. We then have chosen the set of weights that gets the best classification rates among these five. Finally, we classify the test samples considering the weighted total reconstruction error. The produced ROC curves always have a larger AUC than the corresponding versions without weights for all values of $k$ we tested. This is shown in Figure~\ref{auc_vs_k}. The difference is more pronounced for large values of $k$ and becomes small when $k$ diminishes. Still, the AUC of the weighted classifier is always larger, what indicates that the weighted error is a better classification criteria than the non-weighted one. Finally, we plotted the distribution of the weighted total reconstruction error, that can be seen of Figure~\ref{dist_weighted}. This time, we get a clear boundary between errors from samples of different classes.

@@

To understand how the individual errors contribute to the overall performance, we compared two classifiers: one that uses only the reconstructions from the positive PCs and other that uses only the reconstructions from the negative PCs. In our experiments, both performed poorly when compared to the full classifier, indicating that positives and negatives reconstructions are indeed relevant to the classification. The crucial role of the reconstructions from negative PCs is particularly surprising, since non-pedestrian samples are not related and do not contain a specific pattern. Therefore, we conclude that the importance of the negative reconstructions is not on what they characterize, but instead on what they do not characterize. In this way, the negative reconstruction errors contribute to raise the total reconstruction error when the image does not contains a pedestrian, what makes possible a better classification.

We also investigated the relevance of grayscale and edge errors. In other experiment we compared the accuracy of a classifier that uses only grayscale reconstructions and other classifier that uses only edge reconstructions. This time, they performed very well, with a performance close to the that of the total error classifier. This suggest that there is a lot of redundancy between grayscale and edge information. The classifier based only on edge images got better results, indicating that edge information is more relevant to this classifier. In fact, the edge classifier performed better than the total error classifier itself in some cases and, when the results were worse, the accuracy drop was less than 1\%. For this reason, it is possible to speed up the classification time in a significant way by using only edge information and keep the classification accuracy virtually unchanged.

Since our experiments show that each error has a different relevance for classification, we proposed a classifier that uses a weighted version of the total reconstruction error. The problem with weights, though, is how to find the best values. To get an approximation in an automated way, we used a genetic algorithm with a fitness function that aims the best classification rates for the training samples. This classified performed better than the original version in our experiments for all numbers of PCs we tested. % DESCRIBE CLASSIFICATION RATE VS K

Future work include testing this classifier with other pedestrian databases and verifying if the reconstruction approach is suitable to other detection problems in computer vision, such as face detection. We intend to develop the classifier into a complete detection system, capable of locate objects in large pictures.% It is also possible to employ variations of PCA for reconstruction, like the more robust 2D-PCA \cite{li05}.

>>>>

\begin{document}
%
% paper title
% can use linebreaks \\ within to get better formatting as desired
\title{A weighted image reconstruction based on PCA for pedestrian detection}

@@

\iffinal
  \author{
     \IEEEauthorblockN{Guilherme V. Carvalho, Lailson B. Moraes, George D. C. Cavalcanti, Tsang I. R.}
     \IEEEauthorblockA{Centro de Inform\'{a}tica (CIn)\\
      Universidade Federal de Pernambuco (UFPE)\\
      Recife, Pernambuco, Brasil\\
      \texttt{\url{http://cin.ufpe.br/~viisar}}\\
      \texttt{\url{{gvc,lbm4,gdcc,tir}@cin.ufpe.br}}}

@@

\begin{abstract}
Pedestrian detection is a task associated with security and surveillance systems. Its inherently complex nature makes it a hard challenge to develop a detection system. In this article we present an analysis of a pedestrian detection model based on PCA reconstruction errors. We find out that some components of the original classifier are not strictly necessary and can be eliminated to reduce the detection times. We then propose a weighted version of the classifier, that achieves a better classification accuracy.
\end{abstract}

\begin{IEEEkeywords}
pedestrian detection; pattern recognition; principal component analysis
\end{IEEEkeywords}

@@

Pedestrian detection systems have been widely used and developed throughout the computer vision history. Ranging from still image detection to automated car breaking systems, it is becoming an usual task in our lives.

In this paper we describe a pedestrian detection proposed by Malang\'{o}n-Borja and Fuentes \cite{borja09}. It classifies an image based on how much its reconstruction is different from the original one, that is, it classifies based on the image reconstruction error. We analyze how it works and how its original performance can be improved. The proposed improvements enhances the system's accuracy using weights on the detection errors and significantly reduces execution times by decreasing the number of calculations made by the algorithm.

% In this paper we describe a pedestrian detection model proposed by Malag\'{o}n-Borja and Fuentes \cite{borja09}. It classifies an image based on how much its reconstruction is different from the original one; in other words, based on the image reconstruction error. We analyze how it works and how its original performance can be improved, as well as its detection times, by using weights on the reconstruction errors and reducing the number of calculations made by the algorithm, respectively.

This paper is divided as follows. Section 2 describes an image reconstruction technique using PCA. Section 3 describes how to use image reconstruction to classify images. Section 4 presents our experiments and discussions. Finally, in Section 5 conclusions are presented.

@@

The idea of principal component analysis (PCA) seems to have been first proposed in 1901 \cite{sirovich87} and was later developed in many works during the last century. It was popularized in computer vision by the eigenfaces method \cite{turk91}, a face recognition model still relevant and used in current days.

Given a set of samples, PCA yields a set of orthonormal vectors that can be used to linearly project the samples into a new space. This space maximizes the projected samples variance and minimizes their least mean square error, that is, it minimizes the difference between the projected sample and its reconstruction back to the original space \cite{shlens09}. This characteristic makes PCA suitable for data compression and dimensionality reduction. The data can come from many kinds of sources, but in this work we are concerned only with images.

The formulation of PCA is as follows. Consider a set of $m$ images, each with $r$ pixels of width and $c$ pixels of height (i.e. of size $r \times c$ pixels). Each image $I_i$ is represented by a column vector $v_i$ of length $rc$. The set mean $\mu$ and the covariance matrix $C$ are given by

@@

PCA applies eigen-decomposition on $C$ and keeps the $k$ eigenvectors ($1 \leq k \leq rc$) corresponding to the $k$ largest eigenvalues. These eigenvectors are called the \emph{principal components} (PCs) of $C$. It is proven that a projection onto the space defined by these eigenvectors provides the optimal reduced representation of the data, minimizing information loss \cite{shlens09}.

@@

\noindent As the projection is reversible, we can recover the original version of a previously projected image. The \emph{reconstructed image} $u'$ of a projection $p$ is

@@

\noindent This process can be interpreted as a form of lossy compression and decompression: the reconstructed image is generally not equal to the original one, being just an approximation of it. For this reason, each reconstruction has an associated \emph{reconstruction error}, which can be measured by

@@

The more principal components we use to obtain a projection, the less information loss we have, what allows a more accurate reconstruction. Also, the more similar $u$ is to the images that were used to produce $P$, the better the reconstruction should be for a fixed number of eigenvectors.

@@

Based on this fact, Malag'{o}n-Borja and Fuentes \cite{borja09} proposed a classifier for pedestrian detection that uses reconstruction errors as the classification criteria. In addition to the grayscale images, the classifier also uses the corresponding computed edge images, since they characterize better the pedestrian silhouette and eliminate background and clothing variations.

@@

\begin{itemize}
	\item The principal components $P_{gp}$ and the mean $\mu_{gp}$ from the set of pedestrian grayscale images;
	\item The principal components $P_{ep}$ and the mean $\mu_{ep}$ from the set of pedestrian edge images;
	\item The principal components $P_{gn}$ and the mean $\mu_{gn}$ from the set of non-pedestrian grayscale images;
	\item The principal components $P_{en}$ and the mean $\mu_{en}$ from the set of non-pedestrian edge images.
\end{itemize}

In the classification phase, when a new grayscale image has to be classified, the first step is to obtain its edge image and perform four reconstructions: one for each set of PCs. Then, the reconstruction errors must be calculated and combined to produce the total reconstruction error, that is the final classification score. If the total error is greater than or equal to zero, the image is classified as a pedestrian; otherwise, it is classifier as a non-pedestrian. The detailed procedure is described below.

\begin{enumerate}
  \item Obtain the edge image $e$ from grayscale image $g$ using the Sobel operator;
  \item Perform four image reconstructions:
    \begin{enumerate}
      \item $r_{gp} = P_{gp} P_{gp}^T(g-u) + \mu_{gp}$
      \item $r_{ep} = P_{ep} P_{ep}^T(e-u) + \mu_{ep}$
      \item $r_{gn} = P_{gn} P_{gn}^T(g-u) + \mu_{gn}$
      \item $r_{en} = P_{en} P_{en}^T(e-u) + \mu_{en}$
    \end{enumerate}
  \item Calculate the corresponding reconstruction errors:
    \begin{enumerate}
      \item $d_{gp} = |r_{gp} - g|$
      \item $d_{ep} = |r_{ep} - e|$
      \item $d_{gn} = |r_{gp} - g|$
      \item $d_{en} = |r_{en} - e|$
    \end{enumerate}
  \item Compute the \emph{total reconstruction error}, given by
    \begin{equation}
      d_t = d_{gn} + d_{en} - d_{gp} - d_{ep}
      \label{total_error}
    \end{equation}
  \item Classify the image according to:
    \begin{equation}
      \textrm{class}(g) = \left\{
        \begin{array}{l l}
         \textrm{Pedestrian}, & d_t \geq 0\\
         \textrm{Non-pedestrian}, & d_t < 0
				 \label{classification_score}
        \end{array}
      \right.
    \end{equation}

@@

When we analyze how this classifier works, it is easy to understand the role of the PCs generated by the positive samples. As the samples contains the recurrent pattern of the target object (pedestrians in this case), PCA is able to capture it and better reconstruct similar images, generating smaller reconstruction errors. On the other hand, the role of negative PCs is not so clear. Since the negative images contain just random objects and does not represent a specific concept, what pattern is there to capture? We initially thought that the errors of reconstructions obtained from the negative PCs had a negligible importance and therefore it would be possible to eliminate them without impacting the classification accuracy. However, our experiments show that is not the case: it turns out that these errors have a major importance on classification.

We found that the value of the negative PCs is not on the pattern PCA captures, but instead it lies on what pattern it does \emph{not} capture. In this way, the negative reconstruction errors tend to be high for images that contains the target object while for the same images the positive reconstruction errors tend to be small, contributing to a high total error, what classifies the image correctly. For images that does not contains the target object, it is exactly the opposite. Thus, both components of total error (from positive PCs and from negative PCs) work together for classification and that is why it is not worth to construct a classifier that uses only positive reconstruction errors.

However, this does not mean that all the four components have the same relevance to classification. In the reconstruction error classifier, each component contributes equally to the total error, but our experiments show that this relation is not true. Some components have more importance than others and therefore have to be adjusted accordingly. Thus, we propose the \emph{weighted total reconstruction error}, given by

@@

It is also worth noting that the classifier can be made more flexible if we introduce a threshold parameter and compare the classification score (i.e. the total error) agains it, instead of keep it always fixed to zero (as shown in eq.~\ref{classification_score}). In this way, the users can adjust the specificity, depending on their particular application.

@@

The classifier was tested in the context of the pedestrian detection task, that is, given an unlabeled image, the classifier must determine if it depicts or not a pedestrian. To train the classifier, we need a set of pedestrians images and a set of non-pedestrian images. This training phase consists of computing image edges and applying PCA to the four image sets separately, what produces four sets of PCs. They are used to perform the reconstructions on classification time.

The pedestrian images were obtained from the MIT CBCL pedestrian database \cite{cbcl}. It contains 924 color images of pedestrians in frontal or rear views, under different scene conditions. These images were converted to grayscale and had part of the background cropped on the border. The final samples had $45 \times 105$ pixels. This reduces the non-relevant background variation among the images.

@@

To train the classifier, we selected 75\% of the pedestrian images (693 samples) and the remaining images (231 samples) were used  for testing. From the non-pedestrian images, we selected 80\% of the images (4.000 samples) for training the classifier and the remaining images (1.000 samples) were used for testing. This particular scheme was chosen in an empirical way; it yielded the best results among the distributions we tested. The edge images are computed with $x$ and $y$ Sobel filters \cite{gonzalez07}, which we combine to generate a single edge image. To implement the classifier and the auxiliary procedures, we used the MATLAB 2009b environment. The resulting code is available for download at \url{http://github.com/lailsonbm/pca_reconstruction/}.

We start our analysis by investigating the impact of the reconstruction errors from the positive PCs and from the negative PCs to the classification. As discussed before, we initially questioned the role of negative PCs, since they are computed from completely uncorrelated images and do not represent a particular object. However, when we classify test samples using reconstruction errors only from the positive PCs or only from the negative PCs (but still using grayscale and edge images), it becomes clear that neither one can perform a satisfactory classification; both classifiers yield a poor result. But when the errors are combined accordingly to the total reconstruction error, we have a good accuracy. The receiver operating characteristic (ROC) curves of the three classifiers are shown in Figure~\ref{roc_pos_vs_neg} for 100 PCs, i.e., $k = 100$. We get a similar pattern for all values of $k$ that were tested.

\begin{figure}[b!]
\centering
\includegraphics[width=3in]{roc_pos_vs_neg}
\caption{Comparison of positive and negative ROC curves. From left to right: the first curve (green) refers to a classifier that uses the total reconstruction error (eq. \ref{total_error}) and has an AUC of 0.995; the second curve (red) refers to a classifier that uses only the reconstruction errors from negative PCs ($d_{gn}$ and $d_{en}$) and has an AUC of 0.690; and the third curve (blue) refers to a classifier that uses only the errors from positive PCs ($d_{gp}$ and $d_{ep}$) and has an AUC of 0.444. On the three classifiers, only the first 100 principal components are being used, that is, $k=100$. Both positive and negative classifiers yields a poor result, far from the full classifier performance. It is surprising to see that the negative classifier even got a better result than the positive one.}
\label{roc_pos_vs_neg}
\end{figure}

@@

\begin{figure}[t!]
\centering
\includegraphics[width=3in]{roc_gray_vs_edge}
\caption{Comparison of grayscale and edge ROC curves. The green curve refers to a classifier that uses the total reconstruction error (eq. \ref{total_error}) and has an AUC of 0.995; the red curve refers to a classifier that uses only the reconstruction errors from edge PCs ($d_{ep}$ and $d_{en}$) and has an AUC of 0.994; and the blue curve uses only the errors from grayscale PCs ($d_{gp}$ and $d_{gn}$) and has an AUC of 0.985. On the three classifiers, only the first 100 principal components are being used, that is, $k=100$. Note that the scale is different of that from Figure~\ref{roc_pos_vs_neg}; we approximated the upper left corner to better show the small difference between the curves.}
\label{roc_gray_vs_edge}
\end{figure}

Again, we plot the reconstruction error distributions, but now segregating errors from grayscale PCs and from edges PCs. These plots are shown in Figure~\ref{dist_gray_vs_edge}, that also displays a plot of the total reconstruction error. Now, we get a different scenario. All the three plots exhibit a clear boundary that separate samples from different classes. This indicate that there is a lot of redundancy between errors from grayscale PCs and from edge PCs and one of them can be safely removed from the classifier with just a minor impact on the classification accuracy. Our results indicate that the errors from grayscale PCs generally yields a slightly worse result, as shown in the Table~\ref{table_gray_vs_edge}. Therefore, it is possible to compose a classifier only with the edge errors and speed up the classification significantly (since two reconstructions will not have to be computed anymore). In general, such classifier drops the classification accuracy in less than $1\%$ when compared with a classifier that uses the total error. Actually, the classifier based only on the edge errors sometimes performs even better than the full classifier in our experiments (when $k >= 200$), as also shown in Table~\ref{table_gray_vs_edge}.

@@

% What reference use to GA?
These findings point out that, among the four reconstruction errors, some are more relevant to the classification than others. However, on the total reconstruction error (eq. \ref{total_error}) each error contributes equally. Thus, we use the weighted total reconstruction error (eq. \ref{weighted_total_error}) to improve the classification accuracy. The problem with using weights is how to choose them. We can try some values and make adjustments manually, test after test. But we can also employ an optimization algorithm that looks for the values in an automated way. In this work, we use a genetic algorithm~\cite{goldberg89}. It initially chooses a set of random values and uses operations as crossover and mutation to improve them during many iterations. The values are adjusted in relation to a fitness function. In our case, this functions aims to correctly classify the largest number of training samples using a given set of weights. The function can be fine-tuned in order to give more importance in finding more pedestrians despite the false detections or to minimize the number of non-pedestrians classified as pedestrians.

As this algorithm performs a local optimization and therefore is not optimum, we runned it five times for each value of $k$. We then have chosen the set of weights that gets the best classification rates among these five. Finally, we classify the test samples considering the weighted total reconstruction error. The produced ROC curves always have a larger AUC than the corresponding versions without weights for all values of $k$ we tested. This is shown in Figure~\ref{auc_vs_k}. The difference is more pronounced for large values of $k$ and becomes small when $k$ diminishes. Still, the AUC of the weighted classifier is always larger, what indicates that the weighted error is a better classification criteria than the non-weighted one. Finally, we plotted the distribution of the weighted total reconstruction error, that can be seen in Figure~\ref{dist_weighted}. This time, we get a clear boundary between errors from samples of different classes.

@@

To understand how the individual errors contribute to the overall performance, we compared two classifiers: one that uses only the reconstructions from the positive PCs and other that uses only the reconstructions from the negative PCs. In our experiments, both performed poorly when compared to the full classifier, indicating that positives and negatives reconstructions are indeed relevant to the classification. The crucial role of the reconstructions from negative PCs is particularly surprising, since non-pedestrian samples are not related and do not contain a specific pattern. Therefore, we conclude that the importance of the negative reconstructions is not on what they characterize, but instead on what they do not characterize. In this way, the negative reconstruction errors contribute to raise the total reconstruction error when the image does not contain a pedestrian, what makes possible a better classification.

We also investigated the relevance of grayscale and edge errors. In other experiment we compared the accuracy of a classifier that uses only grayscale reconstructions and other classifier that uses only edge reconstructions. This time, they performed very well, with a performance close to that of the total error classifier. This suggest that there is a lot of redundancy between grayscale and edge information. The classifier based only on edge images got better results, indicating that edge information is more relevant to this classifier. In fact, the edge classifier performed better than the total error classifier itself in some cases and, when the results were worse, the accuracy drop was less than 1\%. For this reason, it is possible to speed up the classification time in a significant way by using only edge information and keep the classification accuracy virtually unchanged.

Since our experiments show that each error has a different relevance for classification, we propose a classifier that uses a weighted version of the total reconstruction error. The problem with weights, though, is how to find the best values. To get an approximation in an automated way, we used a genetic algorithm with a fitness function that aims to achieve the best classification rates for the training samples. This classified performed better than the original version in our experiments for all numbers of PCs we tested. % DESCRIBE CLASSIFICATION RATE VS K

% Future work include testing this classifier with other pedestrian databases and verifying if the reconstruction approach is suitable to other detection problems in computer vision, such as face detection. We intend to develop the classifier into a complete detection system, capable of locate objects in large pictures.% It is also possible to employ variations of PCA for reconstruction, like the more robust 2D-PCA \cite{li05}.
